CUBE: An \tcb{Information-optimized parallel} Cosmological $N$-body Algorithm

### 2017-12-14 ms_caf/ms_caf.tex
resolution analysis and bug fix in IC.

\section{Accuracy}\label{s.results}
We run a group of simulations to test the accuracy of {\tt CUBE}. We use same seeds to generate same Gaussian random fields in the initial condition generators of {\tt CUBEP3M} and {\tt CUBE}, and then they produce initial conditions of their own format respectively. Then the main $N$-body codes run their own initial conditions to redshift $z=0$.

@@

\tcb{We label $k=0.2k_{\rm Nyquist}$ as vertical dashed lines, where $k_{\rm Nyquist}$ is the scale of fine mesh grids, and the scale of average particle separations. On this scale, the power spectra are offset from nonlinear predictions by at least 20\%. This error is from the PM algorithm and one has to increase the resolution of the simulation to correct these offsets. We do not plot {\tt CUBEP3M} because we examined that {\tt CUBEP3M} and x2v2 of {\tt CUBE} produce same results. Thus, the differences between {\tt CUBEP3M} and different integer formats of {\tt CUBE} are negligible compared to the error of the PM algorithm.}

\tcb{In the lower parts of these four panels study the cross correlations. This time we in turn compare everything with {\tt CUBEP3M}, to emphasize the decorrelation by different integer formats. Firstly, x2v2 is perfectly correlated and cross correlation of x1v2 is nearly 100\% in all cases. In order to quantify the PM error in terms of decorrelation, we run x2v2 simulations with same initial seeds, but 8 times more particles, and measure the decorrelation caused by coarser resolutions. These results are dashed curves (labeled ``PM'') in the lower parts of these four panels. We conclude that cross correlation of x1v1, in all cases, is not worse than the PM errors.}

\tcb{To summarize from Figure \ref{fig.ccc}, in terms of either power spectrum deviation or cross correlation, the error induced by information optimization (even for x1v1) is lower than the error from the PM algorithm, and we can safely use x1v1 for most of the LSS studies. If one still wants to compare the results when PM already contributes significant error ($k\gtrsim 0.2\ k_{\rm Nyquist}$), errors are from nonlinear (lower $m_p$ and/or lower redshift) and are primarily from compressing the velocity information (x2v1 and x1v1).}

@@

The accuracy of this code is adjustable in that we can choose 1-byte/2-byte integers separately for positions and velocities of particles. In the case of using 2-byte integers for both positions and velocities (``x2v2'', and memory can be 12 bpp), the algorithm gives exact results given by traditional $N$-body algorithms. \tcb{Note that the results are exactly same in that they not only produce same physical statistics of LSS, but also the same error (not physical) from the PM algorithm near Nyquist frequencies. In another word, the positions and velocities of each particle are exact. In practice, we only require that the errors from information optimization is much lower than the errors from PM algorithm. In Figure \ref{fig.ccc} we see that it is the case even for the most memory efficient configuration x1v1. This shows that in most LSS studies, when our scales of interest are smaller than $k\simeq 0.2\ k_{\rm Nyquist}$, six 1-byte fixed point numbers contains sufficient information of every $N$-body particle.}

>>>>

\section{Accuracy}\label{s.results}
We run a group of simulations to test the accuracy of {\tt CUBE}. We use same seeds to generate same Gaussian random fields in the initial condition generators of {\tt CUBEP3M} and {\tt CUBE}, and then they produce initial conditions of their own format respectively. Then the main $N$-body codes run their own initial conditions to redshift $z=0$. \tcb{We use the same force kernels as {\tt CUBEP3M} without PP force. Note that near find grid scales there are possibilities to enhance the force kernel to better match the nonlinear power spectrum predictions, however we use the conservative mode of {\tt CUBEP3M} in this paper. An extended PP force and an unbiased force matching algorithm will be added to {\tt CUBE}. The power spectrum studies are presented in \citet{2013MNRAS.436..540H}, and here we focus on the cross correlations between different integer based methods.}

@@

\tcb{We label $k=0.2\,k_{\rm Nyquist}$ as vertical dashed lines, where $k_{\rm Nyquist}$ is the scale of fine mesh grids, and the scale of average particle separations. On this scale, the power spectra are offset from nonlinear predictions by at least 20\%. This error is from the PM algorithm and one has to increase the resolution of the simulation to correct these offsets. We do not plot {\tt CUBEP3M} because we examined that {\tt CUBEP3M} and x2v2 of {\tt CUBE} produce same results. Thus, the differences between {\tt CUBEP3M} and different integer formats of {\tt CUBE} are negligible compared to the error of the PM algorithm.}

\tcb{In the lower parts of these four panels study the cross correlations. This time we in turn compare everything with {\tt CUBEP3M}, to emphasize the decorrelation by different integer formats. Firstly, x2v2 is perfectly correlated and cross correlation of x1v2 is nearly 100\% in all cases. In order to quantify the PM error in terms of decorrelation, we run x2v2 simulations with same initial conditions, but 8 times more particles and cells, and measure the decorrelation caused by coarser resolutions. These results are dashed curves (labeled ``PM'') in the lower parts of these four panels. We conclude that cross correlation of x1v1, in all cases, is not worse than the PM errors.}

\tcb{To summarize from Figure \ref{fig.ccc}, in terms of either power spectrum deviation or cross correlation, the error induced by information optimization (even for x1v1) is lower than the error from the PM algorithm, and we can safely use x1v1 for most of the LSS studies. If one still wants to compare the results when PM already contributes significant error ($k\gtrsim 0.2\,k_{\rm Nyquist}$), errors are from nonlinear (lower $m_p$ and/or lower redshift) and are primarily from compressing the velocity information (x2v1 and x1v1).}

@@

The accuracy of this code is adjustable in that we can choose 1-byte/2-byte integers separately for positions and velocities of particles. In the case of using 2-byte integers for both positions and velocities (``x2v2'', and memory can be 12 bpp), the algorithm gives exact results given by traditional $N$-body algorithms. \tcb{Note that the results are exactly same in that they not only produce same physical statistics of LSS, but also the same error (not physical) from the PM algorithm near Nyquist frequencies. In another word, the positions and velocities of each particle are exact. In practice, we only require that the errors from information optimization is much lower than the errors from PM algorithm. In Figure \ref{fig.ccc} we see that it is the case even for the most memory efficient configuration x1v1. This shows that in most LSS studies, when our scales of interest are smaller than $k\simeq 0.2\,k_{\rm Nyquist}$, six 1-byte fixed point numbers contains sufficient information of every $N$-body particle.}
